---
title: "Microbiome_Body_Shape_3.29"
author: "Yan Min"
date: "3/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
pkgneeded <- c("phyloseq", "DESeq2", "genefilter", "pheatmap", "dplyr", "gridExtra", "ggplot2", "readxl", "reshape", "cluster", "RColorBrewer", "grid")
sapply(pkgneeded, require, character.only = TRUE)
setwd("~/Documents/WELL_Project/Microbiome/Microbiome_Body_Shape")
```

Fristly load data components: sample data, OTU table, Taxa table, and phylogenetic tree. 
```{r read data}
seqtab <- readRDS("seqtab.rds")
sample <- read.csv("sampleSub_4122018.csv")
sample$X <- NULL
taxatab <- readRDS("taxa.rds")
tax.tab <- as.data.frame(taxatab)
tax.tab$taxa.id <- paste(rep("ID.", nrow(tax.tab)), 1:nrow(tax.tab)) # Assign taxa id to taxa table. use this version of taxa table to match the results. 
phytree <- readRDS("phylo_tree.rds")
df <- read.csv("WELL-China_Firstwave_Metabolic_Disease_Analysis_For_Prediabetes.csv")
df2 <- read.csv("~/Documents/WELL_Project/Microbiome/Microbiome_Body_Shape/WELL.China_Cohort1_20171025.csv", header=TRUE)
id <- sample$id
dfSub <- df[df$id %in% id,]
dfSub.anti <- dfSub[,c("id", "antibiotic", "ratio_fat", "ratio_carbohydrate")]

###Resond to the reviewer's question, examine the menopause 11.19.2018
#df2Sub <- df2[df2$id %in% id, c("id","nomens_why")]

sample <- merge(sample, dfSub.anti, by = "id")
sample$antibiotic <- as.factor(sample$antibiotic)
###Merge the last mens_age to the sample data 11.19.2018
#sample <- merge(sample, df2Sub, by = "id")
sample <- sample[!is.na(sample$antibiotic),] # only keep the complete cases in the sample (1 male was deleted for not having antibiotic information)
### Create another dummy variable indicating menopause status
#sample$meno <- 0
#sample$meno[sample$nomens_why==3] <- 1
rownames(sample) <- sample$Number
seqtab <- seqtab[rownames(seqtab) %in% rownames(sample),] # only keep the complete cases in the otu table
```

```{r}
taxatab["TACGTAGGGGGCAAGCGTTATCCGGATTTACTGGGTGTAAAGGGAGCGTAGACGGCATGGCAAGCCAGATGTGAAAGCCCGGGGCTCAACCCCGGGACTGCATTTGGAACTGTCAGGCTAGAGTGTCGGAGAGGAAAGCGGAATTCCTAGTGTAGCGGTGAAATGCGTAGATATTAGGAGGAACACCAGTGGCGAAGGCGGCTTTCTGGACGATGACTGACGTTGAGGCTCGAAAGCGTGGGGAGCAAACAGG", 6] <- "Blautia" #100%
taxatab["TACGTATGGAGCAAGCGTTATCCGGATTTACTGGGTGTAAAGGGAGTGTAGGTGGCCAGGCAAGTCAGAAGTGAAAGCCCGGGGCTCAACCCCGGGACTGCTTTTGAAACTGCAGGGCTAGAGTGCAGGAGGGGCAAGTGGAATTCCTAGTGTAGCGGTGAAATGCGTAGATATTAGGAGGAACACCAGTGGCGAAGGCGGCTTGCTGGACTGTAACTGACACTGAGGCTCGAAAGCGTGGGGAGCAAACAGG", 3:6] <- c("Bacilli", "Lactobacillales", "Lactobacillaceae", "Lactobacillus") #100%
taxatab["TACGTATGGAGCAAGCGTTATCCGGATTTACTGGGTGTAAAGGGAGTGTAGGTGGCCATGCAAGTCAGAAGTGAAAATCCGGGGCTCAACCCCGGAACTGCTTTTGAAACTGTGAGGCTAGAGTGCAGGAGGGGTGAGTGGAATTCCTAGTGTAGCGGTGAAATGCGTAGATATTAGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGACTGTAACTGACACTGAGGCTCGAAAGCGTGGGGAGCAAACAGG", 5:6] <- c("Eubacteriaceae", "Eubacterium") #99%

```

Construct they phyloseq object.
```{r phyloseq object}
ps <- phyloseq(otu_table(seqtab, taxa_are_rows = FALSE), sample_data(sample), tax_table(taxatab), phy_tree(phytree))
saveRDS(ps, file = "phyloseqObject_9.14.2018.rds")
```

Stratify by gender 
```{r stratification}
ps.f <- subset_samples(ps, gender %in% "Female")
ps.m <- subset_samples(ps, gender %in% "Male")
```

Create the categorical variabbles for android fat mass, gynoid fat mass, and aoi in both sexes seperately.
```{r}
sample_data(ps.f)$Android_FM_Cat <- cut(sample_data(ps.f)$Android_FM, breaks = c(quantile(sample_data(ps.f)$Android_FM, probs = seq(0, 1, by = 0.25))), include.lowest = TRUE)
sample_data(ps.f)$Gynoid_FM_Cat <- cut(sample_data(ps.f)$Gynoid_FM, breaks = c(quantile(sample_data(ps.f)$Gynoid_FM, probs = seq(0, 1, by = 0.25))), include.lowest = TRUE)
sample_data(ps.f)$aoi_Cat <- cut(sample_data(ps.f)$aoi, breaks = c(quantile(sample_data(ps.f)$aoi, probs = seq(0, 1, by = 0.25))), include.lowest = TRUE)
sample_data(ps.f)$Total_FM_Cat <- cut(sample_data(ps.f)$total_fm_ratio, breaks = 4, include.lowest = TRUE)

levels(sample_data(ps.f)$Android_FM_Cat) <- c(1,2,3,4)
levels(sample_data(ps.f)$Gynoid_FM_Cat) <- c(1,2,3,4)
levels(sample_data(ps.f)$aoi_Cat) <- c(1,2,3,4)
levels(sample_data(ps.f)$Total_FM_Cat) <- c(1,2,3,4)

sample_data(ps.m)$Android_FM_Cat <- cut(sample_data(ps.m)$Android_FM, breaks = c(quantile(sample_data(ps.m)$Android_FM, probs = seq(0, 1, by = 0.25))), include.lowest = TRUE)
sample_data(ps.m)$Gynoid_FM_Cat <- cut(sample_data(ps.m)$Gynoid_FM, breaks = c(quantile(sample_data(ps.m)$Gynoid_FM, probs = seq(0, 1, by = 0.25))), include.lowest = TRUE)
sample_data(ps.m)$aoi_Cat <- cut(sample_data(ps.m)$aoi, breaks = c(quantile(sample_data(ps.m)$aoi, probs = seq(0, 1, by = 0.25))), include.lowest = TRUE)
sample_data(ps.m)$Total_FM_Cat <- cut(sample_data(ps.m)$total_fm_ratio, breaks = 4, include.lowest = TRUE)
levels(sample_data(ps.m)$Android_FM_Cat) <- c(1,2,3,4)
levels(sample_data(ps.m)$Gynoid_FM_Cat) <- c(1,2,3,4)
levels(sample_data(ps.m)$aoi_Cat) <- c(1,2,3,4)
levels(sample_data(ps.m)$Total_FM_Cat) <- c(1,2,3,4)
```

Scale the sample data. 
```{r}
sample_data(ps.f) <- as.data.frame(cbind(sample_data(ps.f)[,c(1:8,29:33)], scale(sample_data(ps.f)[,c(9:28)])))
sample_data(ps.m) <- as.data.frame(cbind(sample_data(ps.m)[,c(1:8,29:33)], scale(sample_data(ps.m)[,c(9:28)])))
```

Apply filter to filt out the highly abundant taxa in male and female participants separately. 
Definition of highly abundant: seen more than 5 counts in 7% or more of the samples of each sex.
```{r}
ps.f.filt <- filter_taxa(ps.f, function(x) sum(x>5) > (0.07*length(x)), prune = TRUE) # leaves 336 taxa in 116 female
ps.m.filt <- filter_taxa(ps.m, function (x) sum(x>5) > (0.07*length(x)), prune = TRUE) # leaves 323 taxa in 97 male 
```

6.24.2018 Save ps.f.filt & ps.m.filt for revision analysis (Re Prof.Zhu's questions)
```{r}
#saveRDS(ps.f.filt, file = "ps.f.filt")
#saveRDS(ps.m.filt, file = "ps.m.filt")
```

Prepare DESeq objects (summarized experiment)
```{r}
col.f <- sample_data(ps.f.filt)
col.m <- sample_data(ps.m.filt)
count.f <- t(as.data.frame(otu_table(ps.f.filt)))
count.m <- t(as.data.frame(otu_table(ps.m.filt)))
```

Construct DESeq models.
Firstly test taxa associated with android fat mass (here we use the android fat mass to total fat mass ratio)
```{r}
dds.android.f <- DESeqDataSetFromMatrix(countData = count.f, colData = col.f, design = ~ age + batch.cat + diet_fat + diet_carbohydrate + smoke + alcohol + bmi + antibiotic + Android_FM)
dds.android.m <- DESeqDataSetFromMatrix(countData = count.m, colData = col.m, design = ~ age + batch.cat + diet_fat + diet_carbohydrate + smoke + alcohol + bmi + antibiotic + Android_FM)
```

Estimate size factor & dispersion
```{r}
dds.android.f <- estimateSizeFactors(dds.android.f, type = "poscounts")
dds.android.m <- estimateSizeFactors(dds.android.m, type = "poscounts")

dds.android.f <- estimateDispersions(dds.android.f, fitType = "local")
dds.android.m <- estimateDispersions(dds.android.m, fitType = "local")

#plotDispEsts(dds.android.f)
#plotDispEsts(dds.android.m)
```

Extract the variance stablized data for the Heatmaps.
```{r}
### this part is not changed yet during 6.25 revision!!! Depending on which count numbers will be used for the heatmap
vst.android.f <- getVarianceStabilizedData(dds.android.f)
vst.android.m <- getVarianceStabilizedData(dds.android.m)
#saveRDS(vst.android.f, file = "vst.android.f.rds")
#saveRDS(vst.android.m, file = "vst.android.m.rds")
```

Wald test for android fat.
```{r}
dds.android.f <- nbinomWaldTest(dds.android.f)
dds.android.m <- nbinomWaldTest(dds.android.m)
```

Access the results.
```{r}
res.android.f <- results(dds.android.f, alpha = 0.01, ifcThreadshold = 1)
res.android.m <- results(dds.android.m, alpha = 0.01, ifcThreadshold = 1)
```

Filter the results using padj < 0.01, log2foldchange >1 for each sex.
```{r}
res.android.f.positive <- res.android.f[res.android.f$log2FoldChange > 2 & res.android.f$padj < 0.01 & !is.na(res.android.f$padj),]
res.android.m.positive <- res.android.m[res.android.m$log2FoldChange > 2 & res.android.m$padj < 0.01 & !is.na(res.android.m$padj),]
res.android.f.negative <- res.android.f[res.android.f$log2FoldChange < -2 & res.android.f$padj < 0.01 & !is.na(res.android.f$padj),]
res.android.m.negative <- res.android.m[res.android.m$log2FoldChange < -2 & res.android.m$padj < 0.01 & !is.na(res.android.m$padj),]
```

Match taxa to the taxa table.
```{r}
matchTaxa <- function(result, reference){
  names <- rownames(result)
  taxa <- as.data.frame(reference[rownames(reference) %in% names,])
} # Function to create a taxa table for all the significant taxa. 

otu.android.f.positive <- matchTaxa(res.android.f.positive, tax.tab)
otu.android.m.positive <- matchTaxa(res.android.m.positive, tax.tab)
otu.android.f.negative <- matchTaxa(res.android.f.negative, tax.tab)
otu.android.m.negative <- matchTaxa(res.android.m.negative, tax.tab)
```

Create the result table for paper
```{r}
android.f.positive <- as.data.frame(cbind(otu.android.f.positive,res.android.f.positive[,c(2,6)]))
android.m.positive <- as.data.frame(cbind(otu.android.m.positive,res.android.m.positive[,c(2,6)]))
android.f.positive$taxa <- rownames(android.f.positive)
android.m.positive$taxa <- rownames(android.m.positive)
android.f.negative <- as.data.frame(cbind(otu.android.f.negative,res.android.f.negative[,c(2,6)]))
android.m.negative <- as.data.frame(cbind(otu.android.m.negative,res.android.m.negative[,c(2,6)]))
android.f.negative$taxa <- rownames(android.f.negative)
android.m.negative$taxa <- rownames(android.m.negative)
android.list <- list(a.f.p = android.f.positive, a.f.n = android.f.negative, a.m.p = android.m.positive, a.m.n = android.m.negative)
mapply(write.csv, android.list, file = paste0(names(android.list), ".csv"))
```

There are no pairwise intersect between all 4 android fat result tables. (Female positive & negative; male positive & negative)

Secondly test taxa associated with gynoid fat mass (also use gynoid fat mass to total fat mass ratio)
```{r}
dds.gynoid.f <- DESeqDataSetFromMatrix(countData = count.f, colData = col.f, design = ~ age + batch.cat + diet_fat + diet_carbohydrate + smoke + alcohol + bmi + antibiotic + Gynoid_FM)
dds.gynoid.m <- DESeqDataSetFromMatrix(countData = count.m, colData = col.m, design = ~ age + batch.cat + diet_fat + diet_carbohydrate + smoke + alcohol + bmi + antibiotic + Gynoid_FM)
```

Estimate size factor & dispersion
```{r}
dds.gynoid.f <- estimateSizeFactors(dds.gynoid.f, type = "poscounts")
dds.gynoid.m <- estimateSizeFactors(dds.gynoid.m, type = "poscounts")

dds.gynoid.f <- estimateDispersions(dds.gynoid.f, fitType = "local")
dds.gynoid.m <- estimateDispersions(dds.gynoid.m, fitType = "local")

#plotDispEsts(dds.gynoid.f)
#plotDispEsts(dds.gynoid.m)
```

Extract the variance stablized data for the Heatmaps.
```{r}
### also not changed during 6.25 revision, depending on what to use for the heatmap
vst.gynoid.f <- getVarianceStabilizedData(dds.gynoid.f)
vst.gynoid.m <- getVarianceStabilizedData(dds.gynoid.m)
#saveRDS(vst.gynoid.f, file = "vst.gynoid.f.rds")
#saveRDS(vst.gynoid.m, file = "vst.gynoid.m.rds")
```

Wald test for gynoid fat.
```{r}
dds.gynoid.f <- nbinomWaldTest(dds.gynoid.f)
dds.gynoid.m <- nbinomWaldTest(dds.gynoid.m)
```

Access the results.
```{r}
res.gynoid.f <- results(dds.gynoid.f, alpha = 0.01, ifcThreadshold = 1)
res.gynoid.m <- results(dds.gynoid.m, alpha = 0.01, ifcThreadshold = 1)
```

Filter the results using padj < 0.01, log2foldchange >1 for each sex.
```{r}
res.gynoid.f.positive <- res.gynoid.f[res.gynoid.f$log2FoldChange > 1 & res.gynoid.f$padj < 0.01 & !is.na(res.gynoid.f$padj),]
res.gynoid.m.positive <- res.gynoid.m[res.gynoid.m$log2FoldChange > 1 & res.gynoid.m$padj < 0.01 & !is.na(res.gynoid.m$padj),]
res.gynoid.f.negative <- res.gynoid.f[res.gynoid.f$log2FoldChange < -1 & res.gynoid.f$padj < 0.01 & !is.na(res.gynoid.f$padj),]
res.gynoid.m.negative <- res.gynoid.m[res.gynoid.m$log2FoldChange < -1 & res.gynoid.m$padj < 0.01 & !is.na(res.gynoid.m$padj),]
```

Match taxa to taxa table.
```{r}
otu.gynoid.f.positive <- matchTaxa(res.gynoid.f.positive, tax.tab)
otu.gynoid.m.positive <- matchTaxa(res.gynoid.m.positive, tax.tab)
otu.gynoid.f.negative <- matchTaxa(res.gynoid.f.negative, tax.tab)
otu.gynoid.m.negative <- matchTaxa(res.gynoid.m.negative, tax.tab)
```

Create result table.
```{r}
gynoid.f.positive <- as.data.frame(cbind(otu.gynoid.f.positive,res.gynoid.f.positive[,c(2,6)]))
gynoid.m.positive <- as.data.frame(cbind(otu.gynoid.m.positive,res.gynoid.m.positive[,c(2,6)]))
gynoid.f.positive$taxa <- rownames(gynoid.f.positive)
gynoid.m.positive$taxa <- rownames(gynoid.m.positive)
gynoid.f.negative <- as.data.frame(cbind(otu.gynoid.f.negative,res.gynoid.f.negative[,c(2,6)]))
gynoid.m.negative <- as.data.frame(cbind(otu.gynoid.m.negative,res.gynoid.m.negative[,c(2,6)]))
gynoid.f.negative$taxa <- rownames(gynoid.f.negative)
gynoid.m.negative$taxa <- rownames(gynoid.m.negative)
gynoid.list <- list(g.f.p = gynoid.f.positive, g.f.n = gynoid.f.negative, g.m.p = gynoid.m.positive, g.m.n = gynoid.m.negative)
mapply(write.csv, gynoid.list, file = paste0(names(gynoid.list), ".csv"))
```

Safe the whole session in case R crashes.
```{r}
save.image(file = "Microbiome_BodyShape_WholeSession.rdata")
```

################################################################################

Firstly create the heatmap for the microbiome abundance of the 4 android fat mass quartiles
```{r order android taxa}
ps.f.vst.android <- phyloseq(otu_table(t(vst.android.f), taxa_are_rows = FALSE), sample_data(sample_data(ps.f.filt)), tax_table(taxatab), phy_tree(phytree)) # create a new phyloseq object using variance stablized count data. Used for 6.28 clustering analysis. 
#saveRDS(ps.f.vst.android, "ps.f.vst.android")
ps.m.vst.android <- phyloseq(otu_table(t(vst.android.m), taxa_are_rows = FALSE), sample_data(sample_data(ps.m.filt)), tax_table(taxatab), phy_tree(phytree)) # create a new phyloseq object using vst count data. Used for 6.28 clustering analysis
#saveRDS(ps.m.vst.android, "ps.m.vst.android")
taxa.order.android.f <- names(sort(taxa_sums(ps.f.vst.android), decreasing = TRUE))
taxa.order.android.m <- names(sort(taxa_sums(ps.m.vst.android), decreasing = TRUE))
quartile <- as.character(seq(1:4))
```

```{r create android heatmaps for female}
plot.list.anfroid.f <- list()
for(x in quartile){
  pshm.android.f <- prune_taxa(taxa.order.android.f[1:30], ps.f.vst.android)
  pshm.android.f <- prune_samples(sample_data(pshm.android.f)$Total_FM_Cat == x, pshm.android.f)
  #sample_data(pshm.android.f)$abund <- rowSums(otu_table(pshm.android.f))
  p <- plot_heatmap(pshm.android.f, taxa.label = "Genus", taxa.order = taxa.order.android.f, sample.order = "Android_FM") + ggtitle(paste("Female Total Fat Ratio Bin:", x)) + theme(plot.title = element_text(size = 12), axis.text.x = element_blank(), axis.ticks.x = element_blank()) + coord_fixed() + xlab(paste("Samples Ordered by Android Fat Ratio", "\n(Low-to-High)")) + scale_fill_continuous(low = "white", high = "red3") + theme(legend.title = element_text(size = 8), legend.text = element_text(size = 6),plot.title = element_text(hjust = 0.5)) 
  plot.list.anfroid.f[[x]] = p
}
grid.arrange(grobs = plot.list.anfroid.f[c(2:3)], nrow = 2)

prunedTaxaAF <- otu_table(prune_taxa(taxa.order.android.f[1:30], ps.f.vst.android))
prunedSampleF <- sample_data(prune_taxa(taxa.order.android.f[1:30], ps.f.vst.android))[,c("Android_FM_Cat", "Gynoid_FM_Cat")]
prunedSampleFFull <- sample_data(prune_taxa(taxa.order.android.f[1:30], ps.f.vst.android))
write.csv(prunedTaxaAF, file = "Figure1_Female_Android.csv")
write.csv(prunedSampleF, file = "Figure1_FemaleSample.csv")
write.csv(prunedSampleFFull, file = "Appendix_1-5_FemaleSampleFull.csv")
```

```{r create android heatmap for male}
plot.list.anfroid.m <- list()
for(x in quartile){
  pshm.android.m <- prune_taxa(taxa.order.android.m[1:30], ps.m.vst.android)
  pshm.android.m <- prune_samples(sample_data(pshm.android.m)$Total_FM_Cat == x, pshm.android.m)
  sample_data(pshm.android.m)$abund <- rowSums(otu_table(pshm.android.m))
  p <- plot_heatmap(pshm.android.m, taxa.label = "Genus", taxa.order = taxa.order.android.m, sample.order = "Android_FM") + ggtitle(paste("Male Total Fat Ratio Bin:", x)) + theme(plot.title = element_text(size = 12), axis.text.x = element_blank(), axis.ticks.x = element_blank()) + coord_fixed() + xlab(paste("Samples Ordered by Anrdoid Fat Ratio", "\n(Low-to-High)")) + scale_fill_continuous(low = "white", high = "red3") + theme(legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(hjust = 0.5)) 
  plot.list.anfroid.m[[x]] = p
}
grid.arrange(grobs = plot.list.anfroid.m[c(2:3)], nrow = 2)

prunedTaxaAM <- otu_table(prune_taxa(taxa.order.android.m[1:30], ps.m.vst.android))
prunedSampleM <- sample_data(prune_taxa(taxa.order.android.m[1:30], ps.m.vst.android))[,c("Android_FM_Cat", "Gynoid_FM_Cat")]
prunedSampleMFull <- sample_data(prune_taxa(taxa.order.android.m[1:30], ps.m.vst.android))
write.csv(prunedTaxaAM, file = "Figure1_Male_Android.csv")
write.csv(prunedSampleM, file = "Figure1_MaleSample.csv")
write.csv(prunedSampleMFull, file = "Appendix_1-5_MaleSampleFull.csv")
```

Then create the heatmap for the microbiome abundance of the 4 gynoid fat mass quartiles
```{r order gynoid taxa}
ps.f.vst.gynoid <- phyloseq(otu_table(t(vst.gynoid.f), taxa_are_rows = FALSE), sample_data(sample_data(ps.f.filt)), tax_table(taxatab), phy_tree(phytree)) # create a new phyloseq object using variance stablized count data. Used for 6.28 clustering analysis. 
saveRDS(ps.f.vst.gynoid, "ps.f.vst.gynoid")
ps.m.vst.gynoid <- phyloseq(otu_table(t(vst.gynoid.m), taxa_are_rows = FALSE), sample_data(sample_data(ps.m.filt)), tax_table(taxatab), phy_tree(phytree)) # create a new phyloseq object using variance stablized count data. Used for 6.28 clustering analysis. 
saveRDS(ps.m.vst.gynoid, "ps.m.vst.gynoid")
taxa.order.gynoid.f <- names(sort(taxa_sums(ps.f.vst.gynoid), decreasing = TRUE))
taxa.order.gynoid.m <- names(sort(taxa_sums(ps.m.vst.gynoid), decreasing = TRUE))
```

```{r create gynoid fat heatmaps for female}
plot.list.gynoid.f <- list()
for(x in quartile){
  pshm.gynoid.f <- prune_taxa(taxa.order.gynoid.f[1:30], ps.f.vst.gynoid)
  pshm.gynoid.f <- prune_samples(sample_data(pshm.gynoid.f)$Total_FM_Cat == x, pshm.gynoid.f)
  sample_data(pshm.gynoid.f)$abund <- rowSums(otu_table(pshm.gynoid.f))
  p <- plot_heatmap(pshm.gynoid.f, taxa.label = "Genus", taxa.order = taxa.order.gynoid.f, sample.order = "Gynoid_FM") + ggtitle(paste("Female Total Fat Ratio Bin:", x)) + theme(plot.title = element_text(size = 12), axis.text.x = element_blank(), axis.ticks.x = element_blank()) + coord_fixed() + xlab(paste("Samples Ordered by Gynoid Fat Ratio", "\n(Low-to-High)")) + scale_fill_continuous(low = "white", high = "red3") + theme(legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(hjust = 0.5)) 
  plot.list.gynoid.f[[x]] = p
}
grid.arrange(grobs = plot.list.gynoid.f[c(2:3)], nrow = 2)

prunedTaxaGF <- otu_table(prune_taxa(taxa.order.gynoid.f[1:30], ps.f.vst.gynoid))
write.csv(prunedTaxaGF, file = "Figure1_Female_Gynoid.csv")
```

```{r create gynoid heatmap for male}
plot.list.gynoid.m <- list()
for(x in quartile){
  pshm.gynoid.m <- prune_taxa(taxa.order.gynoid.m[1:30], ps.m.vst.gynoid)
  pshm.gynoid.m <- prune_samples(sample_data(pshm.gynoid.m)$Total_FM_Cat == x, pshm.gynoid.m)
  sample_data(pshm.gynoid.m)$abund <- rowSums(otu_table(pshm.gynoid.m))
  p <- plot_heatmap(pshm.gynoid.m, taxa.label = "Genus", taxa.order = taxa.order.gynoid.m, sample.order = "Gynoid_FM") + ggtitle(paste("Male Gynoid Fat Mass Bin:", x)) + theme(plot.title = element_text(size = 12), axis.text.x = element_blank(), axis.ticks.x = element_blank()) + coord_fixed() + xlab(paste("Samples Ordered by Gynoid Fat Ratio", "\n(Low-to-High)")) + scale_fill_continuous(low = "white", high = "red3") + theme(legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(hjust = 0.5)) #axis.text.x = element_blank(), 
  plot.list.gynoid.m[[x]] = p
}
grid.arrange(grobs = plot.list.gynoid.m[c(2:3)], nrow = 2)

prunedTaxaGM <- otu_table(prune_taxa(taxa.order.gynoid.m[1:30], ps.m.vst.gynoid))
write.csv(prunedTaxaGM, file = "Figure1_Male_Gynoid.csv")
```

################################################################################
Merge result table, create two new variables:
effect: indicates whether the taxa is protective or harmful, positively related to android fat or negatively related to gynoid fat are considered harmful, coded as 1; negatively related to android fat or positively related to gynoid fat are considered protective, coded as o. 
sex: detected in female coded as 0, male coded as 1
```{r merge results}
merge.result.f <- rbind(android.f.positive, android.f.negative, gynoid.f.positive, gynoid.f.negative)
merge.result.f$effect[merge.result.f$taxa %in% c(rownames(android.f.positive), rownames(gynoid.f.negative))] <- 1
merge.result.f$effect[is.na(merge.result.f$effect)] <- 0
merge.result.f$log2FoldChange <- abs(merge.result.f$log2FoldChange)
merge.result.f$Genus <- as.character(merge.result.f$Genus)
merge.result.f[5,6] <- "Sporobacter" #94%
merge.result.f[20,6] <- "Sporobacter" #94%

write.csv(merge.result.f, file = "merge.result.f.csv")

merge.result.m <- rbind(android.m.positive, android.m.negative, gynoid.m.positive, gynoid.m.negative)
merge.result.m$effect[merge.result.m$taxa %in% c(rownames(android.m.positive), rownames(gynoid.m.negative))] <- 1
merge.result.m$effect[is.na(merge.result.m$effect)] <- 0
merge.result.m$log2FoldChange <- abs(merge.result.m$log2FoldChange)
merge.result.m$Family <- as.character(merge.result.m$Family)
merge.result.m$Genus <- as.character(merge.result.m$Genus)
merge.result.m[1,6] <- "Holdemanella" #97%
merge.result.m[4,6] <- "Absiella" #97%
merge.result.m[5,6] <- "Haemophilus" #100%
merge.result.m[9,c(5,6)] <- c("Clostridiaceae", "Clostridium") #92%
merge.result.m[11,6] <- "Senegalimassilia" #100%
merge.result.m[17,6] <- "Lachnoclostridium" #96%
merge.result.m[24,c(5,6)] <- c("Clostridiales Family XIII. Incertae Sedis", "Ihubacter") #97%

write.csv(merge.result.m, file = "merge.result.m.csv")
```

################################################################################
Finally, generate a demographic table for the paper
```{r}
tableRate <- function(dataframe, varname1, varname2) {
  p <- table(dataframe[,varname1], dataframe[,varname2])
  p <- as.data.frame(p)
  p <- cast(p, Var1 ~ Var2, value = "Freq") # note that the function only support varname2 as a categorical variable with 2 levels 
  colnames(p) <- c(varname1, "male", "female")
  p$total <- p$male + p$female
  p$mper <- p$male/sum(p$male)
  p$fper <- p$female/sum(p$female)
  p$totalper <- p$total/sum(p$total)
  return(p)
}
```

```{r}
tableRate(dfSub, "agecat2", "gender_it")
tableRate(dfSub, "marriage", "gender_it")
tableRate(dfSub, "edu", "gender_it")
tableRate(dfSub, "bmicat2", "gender_it")
tableRate(dfSub, "whrcat", "gender_it")
tableRate(dfSub, "smoke", "gender_it")
tableRate(dfSub, "alcohol", "gender_it")
tableRate(dfSub, "antibiotic", "gender_it")
tableRate(dfSub, "nutri_suppl", "gender_it")
tableRate(dfSub, "meta", "gender_it")
```

Get the correlation coeffcients and variance of android and gynoid fat ratio by gender.
```{r}
fsam <- as.data.frame(sample_data(ps.f.filt))
msam <- as.data.frame(sample_data(ps.m.filt))
```

```{r}
cor(fsam$Android_FM, fsam$Gynoid_FM)
cor(msam$Android_FM, msam$Gynoid_FM)
```

```{r}
originfsam <- sample[sample$gender == "Female",]
originmsam <- sample[sample$gender == "Male",]
sd(originfsam$Android_FM)
sd(originmsam$Android_FM)
sd(originfsam$Gynoid_FM)
sd(originmsam$Gynoid_FM)
```

#### Investigate menopause in women and taxa associated with Android_FM
```{r}
dds.android.f <- DESeqDataSetFromMatrix(countData = count.f, colData = col.f, design = ~ age + batch.cat + diet_fat + diet_carbohydrate + smoke + alcohol + bmi + antibiotic + meno + Android_FM)
dds.android.f <- estimateSizeFactors(dds.android.f, type = "poscounts")
dds.android.f <- estimateDispersions(dds.android.f, fitType = "local")
dds.android.f <- nbinomWaldTest(dds.android.f)
```

```{r}
dds.gynoid.f <- DESeqDataSetFromMatrix(countData = count.f, colData = col.f, design = ~ age + batch.cat + diet_fat + diet_carbohydrate + smoke + alcohol + bmi + antibiotic + Gynoid_FM)
dds.gynoid.f <- estimateSizeFactors(dds.gynoid.f, type = "poscounts")
dds.gynoid.f <- estimateDispersions(dds.gynoid.f, fitType = "local")
dds.gynoid.f <- nbinomWaldTest(dds.gynoid.f)
```

```{r}
res.android.f <- results(dds.android.f, alpha = 0.01)
```

------------- Save all the datasets for figure generation -----------
```{r}
setwd("~/Documents/WELL_Project/Microbiome/Microbiome_Body_Shape")
load("Microbiome_BodyShape_WholeSession.rdata")
```
